### 2021-08-31 presentation/presentation.tex
change color theme

% $Header$

\documentclass[aspectratio=169]{beamer}

% This file is a solution template for:

% - Talk at a conference/colloquium.
% - Talk length is about 20min.
% - Style is ornate.

% Copyright 2004 by Till Tantau <tantau@users.sourceforge.net>.
%
% In principle, this file can be redistributed and/or modified under
% the terms of the GNU Public License, version 2.
%
% However, this file is supposed to be a template to be modified
% for your own needs. For this reason, if you use this file as a
% template and not specifically distribute it as part of a another
% package/program, I grant the extra permission to freely copy and
% modify this file as you see fit and even to delete this copyright
% notice.

\mode<presentation>
{
  \usetheme{Luebeck}
  % \usetheme{Madrid}
  % \usetheme{Warsaw}
  % \usetheme{Copenhagen}
  % \usecolortheme{beaver}
  % \usecolortheme{rose}
  % or ...
  % \usecolortheme{beaver}
  \useoutertheme{split}
  \useinnertheme{circles}

  \setbeamercolor{palette primary}{bg=red!50!black!80}
  \setbeamercolor{palette secondary}{bg=red!50!black!70}
  \setbeamercolor{palette tertiary}{bg=gray!90!black!50}
  \setbeamercolor{palette quaternary}{bg=gray!90!black!100}

  \setbeamercolor{item}{fg=red!50!black!80}
  \setbeamercolor{section in toc}{fg=red!50!black!80}

  % \setbeamercolor{author in head/foot}{bg=grey!60}
  \setbeamercovered{transparent}
\setbeamertemplate{navigation symbols}{}
\makeatletter
\setbeamertemplate{page number in head/foot}[appendixframenumber]
\setbeamertemplate{headline}
{%
  \leavevmode%
  \hbox{%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.25ex,dp=1ex,center]{author in head/foot}%
    \centering
    \usebeamerfont{author in head/foot}\insertsection{}
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.25ex,dp=1ex,right]{title in head/foot}%
    \centering
    \insertsubsection{}
  \end{beamercolorbox}%
  }%
  \vskip0pt%
}
\setbeamertemplate{footline}
{%
  \leavevmode%
  \hbox{%
  \begin{beamercolorbox}[wd=.1\paperwidth,ht=2.25ex,dp=1ex,center]{date in head/foot}%
    \centering
    \usebeamerfont{date in head/foot}\usebeamercolor[fg]{page number in head/foot}\usebeamerfont{page number in head/foot}\usebeamertemplate{page number in head/foot}
    % \hspace*{2ex}
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.25\paperwidth,ht=2.25ex,dp=1ex,center]{author in head/foot}%
    \centering
    \usebeamerfont{author in head/foot}\insertshortauthor{}
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.65\paperwidth,ht=2.25ex,dp=1ex,right]{title in head/foot}%
    \centering
    \insertshorttitle{}
  \end{beamercolorbox}%
  }%
  \vskip0pt%
}
\makeatother

  % or whatever (possibly just delete it)
}

\usepackage[english]{babel}
% or whatever

\usepackage[utf8]{inputenc}
% or whatever

% \usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\title[Detection and Mitigation of Corrupted Information in DMPC Based on Resource Allocation] % (optional, use only with long paper titles)
{Detection and Mitigation of Corrupted Information in Distributed Model Predictive Control Based on Resource Allocation}

% \subtitle
% {Include Only If Paper Has a Subtitle}

\author[Nogueira, Bourdais, Guéguen] % (optional, use only with lots of authors)
{R. A. Nogueira \and R. Bourdais \and H. Guéguen\\
\texttt{\{rafael-accacio.nogueira,romain.bourdais,herve.gueguen\}\ at \ centralesupelec.fr}}
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[IETR --- CentraleSupélec] % (optional, but mostly needed)
{
  % \inst{1}%
  AUT Department
  \\IETR --- CentraleSupélec
  % \\[10pt]\\\includegraphics[width=1.5cm]{../img/logos/supelec.jpeg}
}
% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

\date[SysTol'21] % (optional, should be abbreviation of conference name)
{5th International Conference on Control and Fault-Tolerant Systems, 2021
                      \includegraphics[width=2cm]{../img/qrPresentation.png}
}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

\subject{}
% This is only inserted into the PDF information catalog. Can be left
% out.

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

\pgfdeclareimage[height=.8cm]{university-logo}{../img/logos/supelec.jpeg}
\logo{\pgfuseimage{university-logo}}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
\AtBeginSubsection[]
{
  \begin{frame}<beamer>{Outline}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

%\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}[plain]
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

% Structuring a talk is a difficult task and the following structure
% may not be suitable. Here are some rules that apply for this
% solution:

% - Exactly two or three sections (other than the summary).
% - At *most* three subsections per section.
% - Talk about 30s to 2min per frame. So there should be between about
%   15 and 30 frames, all told.

% - A conference audience is likely to know very little of what you
%   are going to talk about. So *simplify*!
% - In a 20min talk, getting the main ideas across is hard
%   enough. Leave out details, even if it means being less precise than
%   you think necessary.
% - If you omit details that are vital to the proof/implementation,
%   just say so once. Everybody will be happy with that.

\section{Motivation}

\subsection{The Basic Problem That We Studied}

\begin{frame}{Make Titles Informative. Use Uppercase Letters.}{Subtitles are optional.}
  % - A title should summarize the slide in an understandable fashion
  %   for anyone how does not follow everything on the slide itself.

  \begin{itemize}
  \item
    Use \texttt{itemize} a lot.
  \item
    Use very short sentences or short phrases.
  \end{itemize}
\end{frame}

\begin{frame}{Make Titles Informative.}

  You can create overlays\dots
  \begin{itemize}
  \item using the \texttt{pause} command:
    \begin{itemize}
    \item
      First item.
      \pause
    \item
      Second item.
    \end{itemize}
  \item
    using overlay specifications:
    \begin{itemize}
    \item<3->
      First item.
    \item<4->
      Second item.
    \end{itemize}
  \item
    using the general \texttt{uncover} command:
    \begin{itemize}
      \uncover<5->{\item
        First item.}
      \uncover<6->{\item
        Second item.}
    \end{itemize}
  \end{itemize}
\end{frame}

\subsection{Previous Work}

\begin{frame}{Make Titles Informative.}
\end{frame}

\begin{frame}{Make Titles Informative.}
\end{frame}

\section{Our Results/Contribution}

\subsection{Main Results}

\begin{frame}{Make Titles Informative.}
\end{frame}

\begin{frame}{Make Titles Informative.}
\end{frame}

\begin{frame}{Make Titles Informative.}
\end{frame}

\subsection{Basic Ideas for Proofs/Implementation}

\begin{frame}{Make Titles Informative.}
\end{frame}

\begin{frame}{Make Titles Informative.}
\end{frame}

\begin{frame}{Make Titles Informative.}
\end{frame}

\section*{Summary}

\begin{frame}{Summary}
  \begin{minipage}{.6\textwidth}
    % Keep the summary *very short*.
    \begin{enumerate}
      \item
            Resource allocation based DMPC is vulnerable to attacks.
      \item
            Sub-problems structure has time invariant parameters.
      \item
            Attack can be estimated using these parameters.
    \end{enumerate}

    % The following outlook is optional.
    \vskip0pt plus.5fill
    \begin{itemize}
      \item
            Outlook
            \begin{itemize}
              \item
                    Inequality Constraints yield Hybrid behavior
              \item
                    Non-linear attack model
            \end{itemize}
      \item Repository
\vspace{-.4cm}

                    \begin{minipage}{.6\textwidth}
                      \url{https://github.com/Accacio/SysTol-21}
                    \end{minipage}
                    \begin{minipage}{.2\textwidth}
                      \includegraphics[width=2cm]{../img/qrRepo.png}
                    \end{minipage}
    \end{itemize}
  \end{minipage}
\end{frame}

\nocite{VelardeEtAl2017a}
% All of the following is optional and typically not needed.
\appendix
\section<presentation>*{\appendixname}
\subsection<presentation>*{For Further Reading}

\begin{frame}[allowframebreaks]
\bibliographystyle{IEEEtran}
\bibliography{../article/bibliography}
  \frametitle<presentation>{For Further Reading}

  \begin{thebibliography}{10}

  \beamertemplatebookbibitems
  % Start with overview books.
\bibitem{MaestreEtAl2014}
J.~M. Maestre, R.~R. Negenborn \emph{et~al.}, \emph{{D}istributed {M}odel
  {P}redictive {C}ontrol made easy}.\hskip 1em plus 0.5em minus 0.4em\relax
  Springer, 2014, vol.~69.

  \bibitem{Author1990}
    A.~Author.
    \newblock {\em Handbook of Everything}.
    \newblock Some Press, 1990.

  \beamertemplatearticlebibitems
  % Followed by interesting articles. Keep the list short.

  \bibitem{Someone2000}
    S.~Someone.
    \newblock On this and that.
    \newblock {\em Journal of This and That}, 2(1):50--100,
    2000.
  \end{thebibliography}
\end{frame}

\end{document}