### 2021-09-07 img/mpcState10.tex

% This file was created by matlab2tikz.
%
\definecolor{mycolor1}{rgb}{0.47000,0.12000,0.22000}%
%
\begin{tikzpicture}

\begin{axis}[%
width=0.959\fwidth,
height=\fheight,
at={(0\fwidth,0\fheight)},
scale only axis,
xmin=1,
xmax=10,
ymin=18,
ymax=22,
ylabel style={font=\color{white!15!black}},
ylabel={$\text{Temperature (}^\text{o}\text{C)}$},
axis background/.style={fill=white},
title style={font=\bfseries},
title={Temperature in Room},
legend style={at={(0.03,0.97)}, anchor=north west, legend cell align=left, align=left, draw=white!15!black}
]
\addplot[const plot, color=gray, line width=1.5pt] table[row sep=crcr] {%
1	20\\
2	20\\
3	20\\
4	20\\
5	20\\
6	21\\
7	21\\
8	21\\
9	21\\
10	21\\
};
\addlegendentry{Setpoint}

\addplot [color=mycolor1, only marks, mark=triangle*, mark options={solid, fill=mycolor1, mycolor1}]
  table[row sep=crcr]{%
1	18.3\\
2	19.2922432972485\\
3	19.7421558473231\\
4	19.9080020336521\\
5	19.9691994260194\\
6	19.991835840366\\
7	20.6315685911097\\
8	20.8673492565163\\
9	20.9543208818143\\
10	20.9864637751245\\
};
\addlegendentry{State}

\addplot [color=white!50!mycolor1, line width=1.0pt, only marks, mark=triangle, mark options={solid, white!50!mycolor1}]
  table[row sep=crcr]{%
11	20.9349764035864\\
12	20.8913786496969\\
13	20.8170082544264\\
14	20.6458633361213\\
15	20.225626806513\\
};
\addlegendentry{Prediction}

\addplot[area legend, draw=none, fill=mycolor1, fill opacity=0.1, forget plot]
table[row sep=crcr] {%
x	y\\
11	18\\
11	22\\
15	22\\
15	18\\
}--cycle;
\end{axis}
\end{tikzpicture}%